The Etiology of Cybersecurity

### 2020-07-20 reports/paper_0/main.tex
fixed Luca's comments on intro of sec 2 (literature review)

\begin{abstract}
	The objective of this research is the development of a theory that
	determines (all and only) the possible insecurity and security
	configurations of any abstract system.  Our hypothesis is that errors
	are at the bottom of the causality chain that leads to cybersecurity
	attacks. Based on this hypothesis, we formulate a scientific
	(falsifiable by experiments) theory that predicts all the errors that lead to
	the presence of cybersecurity weaknesses in the architectural design of
	an abstract system.  Our theory is based on a correlation between the
	epistemological concepts of knowledge, beliefs, and information with
	the engineering concepts of requirements, functional architectures, and
	channels.  Our theory allows the definition of the security of a system
	as a mathematical formula correlating the security and insecurity
	configurations of a system.  We implemented a prototype cybersecurity
	risk assessment tool that, based on our theory, predicts weaknesses in
	a UML model of a (cyber-physical) system.
\end{abstract}

@@

If we consider a natural (real-world) phenomenon, such as gravity, we can think
of a scientific theory as an \emph{abstract} mathematical formula (or a set of
formulas) that defines axiomatically a certain (natural) phenomenon in a
specific algebraic structure. While such a theory can be proved to be correct
within the mathematical setting it has been conceived in (e.g. with respect to
other axioms of the algebraic structure where it has been postulated, or with
respect to other background theories), no proof can be carried out to verify
that the theory properly abstracts the phenomenon it describes (i.e. that the
theory is a sound or complete approximation of the real phenomenon).
Similarly, the results of the application of a scientific theory (i.e. its
predictions) can be verified in the abstract mathematical structure but how can
we be sure that the theory correctly characterizes the phenomenon it describes?

Popper's famous ``demarcation principle'' draws a line between scientific and
pseudo-scientific theories by proposing that \emph{any} scientific theory can
be considered correct as long as no empirical experiment falsify it. For
example, the theory of gravity, in its mathematical setting, can predict how
objects attract each other, and those predictions can be tested by empirical
experiment. If the theory correctly abstracts the phenomenon, no empirical
experiment should falsify the predictions of that theory.

In \autocite{Herley2016unfalsifiability}, Cormac Herley explores what he calls
``an asymmetry in computer security'', which he defines as follows: ``Things
can be declared insecure by observation, but not the reverse. There is no
observation that allows us to declare an arbitrary system or technique
secure''. His intuition is that there is no scientific theory that can predict
the security of a system, nor a theory that can predict all possible
insecurities of a system (which, by negation would become a theory on
security).  Herley then uses this argument to show that ``claims that any
measure is necessary for security are empirically unfalsifiable''. An example
is whether or not password masking makes a system secure. Password masking
prevents a complete leakage of the password and a system in which password
masking is implemented is believed to be secure because an attacker won't be
seeing the password by ``shoulder surging''.  However, this reasoning is
circular and it is always true that a system in which password masking is
implemented is a system in which a password is masked. No experiment can ever
falsify this statement.  Given that any theory that is not falsifiable by an
empirical experiment is well known\footnote{``A theory which is not refutable
by any conceivable event is nonscientific. Irrefutability is not a virtue of a
theory (as people often think) but a vice.'' -- Karl Popper, Conjectures and
Refutations~\autocite{popper1962conjectures}.} to be nonscientific (i.e.
unfalsifiability is a fallacy of a theory), Herley concludes that there is no
scientific theory of cybersecurity; which means that cybersecurity lays in the
realm of pseudo-sciences~\autocite{Herley2016usenixvideo}. Herley, e.g.  in
\autocite{Herley2017justifying}, discusses the implications of a nonscientific
approach to cybersecurity, and highlights the tremendous impact on all the
scientific research and engineering of systems. While the criticism is
investigated in \autocite{Herley2016unfalsifiability}, no solution is provided
nor envisioned.

@@

\section{Literature Review}\label{sec:literature}
\fixnote{Luca}{I think that this section needs thorough rewriting. First of all, you are trying to say too many things, many of which are not really relevant for what you want to say. Second, you are not just reviewing but also introducing some new things that are part of your contributions and this is quite confusing. Third, there are some claims that, in my view, are wrong or at best misleading. Shorten, simplify, aiming to distill really only what is relevant. As it is now, the text is quite confused and confusing.}
\begin{figure*}[t]
	\centering
	\includegraphics[width=.8\textwidth]{safety-security_3.pdf}
	\caption{Overview of keywords related to security and safety}
	\label{fig:safety-security}
\end{figure*}

In most of the natural languages the concepts of safety
and security are not syntactically differentiated and both terms (safety and
security) are expressed by the same word, e.g. sicurezza in Italian.  A
semantic distinction between safety and security is correlated to a
belief\footnote{A belief has to be intended as a proposition which is supposed
to be true by the majority of people in our society, without a scientific
underlying theory but based on partial empirical evidences or inductive reasoning based on partial empirical evidences.} that
safety deals with \emph{accidents} (i.e. an unfortunate incident) posed by the
natural environment (e.g. natural events such as wearing of hardware
components), whereas security deals with \emph{incidents} posed by mankind (e.g.
attackers and bugs). The fundamental difference between nature and mankind (and,
in turn, between safety and cybersecurity\fixnote{Luca}{You spoke of security before, which is much wieder than cybersecurity. This is something you have to be very careful with... political scientists would correctly claim that security is much more than just cyber and that their term is the correct one to contrappose to safety}) is believed to be on the different
intents (incidents are intentional, whereas accidents are not) of the causes
that generates a threat. Namely, nature is believed not to have malicious
intents (but unfortunate causes-effects), whereas threats generated by mankind are
malicious.\footnote{Of course, logical flaws or bugs may be
introduced by other means (e.g. ignorance) without explicit malicious intents,
but the exploitation of those flaws is considered (for now, and detailed
afterwards in the article) malicious, and thus we consider any vulnerability to
be malicious even if it is due to the lack of skills.}.\fixnote{Luca}{Secondo me tutta questa prima discussione \`e alquanto inutile e si pu\`o riassumere in un'unica frase.}
This conclusion seems to be against a general formulation of a cybersecurity theory;
how can we define a theory that predicts what a human being will do?
Cybersecurity attacks seem to be related to the creativity of the attacker
and thus unpredictable. Currently, our understanding of cybersecurity is
stored into a network of databases of weaknesses (e.g.
CWE~\autocite{MITRE2020CWEresearch}), vulnerabilities (e.g. CVE~\autocite{CVE},
NVD~\autocite{NIST2020NVD}),
and attacks (e.g. CAPEC~\autocite{MITRE2020CAPEC}, ATT\&CK~\autocite{MITRE2020ATTACK}).\fixnote{Luca}{Come detto, sono fortemente in disaccordo con questa frase. Parte del nostro understanding \`e l\`i, ma non tutto. Tutto il lavoro teorica e formale non \`e contenuto l\`i. Mi sembra tra l'altro un'affermazione alquanto inutilmente provocatoria} Systems are
nowadays tested against 
known attacks, ignoring that considering a system secure
if no known attack is possible is against the scientific method.\fixnote{Luca}{No, non \`e una scelta ma \`e semplicemente ignoranza di chi fa i test. Chi fa i test seriamente questo lo sa benissimo. E non solo perch\'e \`e contro lo scientific methos ma anche perch\'e testing it not proving}
As an example, if we have a system which is secure against all of the
vulnerabilities in the CVE we often just need to wait for a month to have 1000+ more
vulnerabilities to test \autocite{NVD2020dashboard}. If even just one of those new vulnerabilities
affects our system, that falsifies the hypothesis that a
system is secure if no known attacks apply.\fixnote{Luca}{But, again, the hypothesis was utterly wrong from the start, even without considering the scientific method. No person who is really knowledgeable of testing would ever claim that testing shows the absence of all attacks. Maybe this is what is done in industry, but then industry is simply wrong (it would not be the first time) and people in academia have been saying so for ages. It is not you argueing it.. I believe that building the argument as you do is wrong. I would suggest: (1) Industry uses testing believing that it is enough. (2) It is well know that testing cannot show the absence of attacks. (3) The scientific method makes this even more evident and in fact we show ...} 
We argue that this ``implement and test'' approach should be changed.
Even tough the field 
of security protocols is still lacking of a general cybersecurity theory,
the secure-by-design approach they apply can be considered as going
into the opposite direction of the ``implement and test'' approach.\fixnote{Luca}{Questa ultima frase \`e appesa qui senza una vera chiara connessione a quanto detto finora.}

>>>>

\begin{abstract}
	The objective of this research is the development of a theory that
	determines (all and only) the possible insecurity and security
	configurations of any abstract system.  Our hypothesis is that errors
	are at the bottom of the causality chain that leads to cybersecurity
	attacks. Based on this hypothesis, we formulate a scientific
	(falsifiable by experiments) theory that predicts all the errors that lead to
	the presence of cybersecurity weaknesses in the architectural design of
	an abstract system.  Our theory is based on a correlation between the
	epistemological concepts of knowledge, beliefs, and information with
	the engineering concepts of requirements, functional architectures, and
	channels.  Ultimately, our theory allows the definition of cybersecurity of a system
	as a mathematical formula.  We implemented a prototype cybersecurity
	risk assessment tool that, based on our theory, predicts weaknesses in
	a UML model of a (cyber-physical) system.
\end{abstract}

@@

If we consider a natural (real-world) phenomenon, such as gravity, we can think
of a scientific theory as an \emph{abstract} mathematical formula (or a set of
formulas) that defines axiomatically a certain (natural) phenomenon in a
specific algebraic structure. While such a theory can be proved to be correct
within the mathematical setting it has been conceived in (e.g. with respect to
other axioms of the algebraic structure where it has been postulated, or with
respect to other background theories), no proof can be carried out to verify
that the theory properly abstracts the phenomenon it describes (i.e. that the
theory is a sound or complete approximation of the real phenomenon).
Similarly, the results of the application of a scientific theory (i.e. its
predictions) can be verified in the abstract mathematical structure but how can
we be sure that the theory correctly characterizes the phenomenon it describes? 
Popper's famous ``demarcation principle'' draws a line between scientific and
pseudo-scientific theories by proposing that \emph{any} scientific theory can
be considered correct as long as no empirical experiment falsify it. For
example, the theory of gravity, in its mathematical setting, can predict how
objects attract each other, and those predictions can be tested by empirical
experiment. If the theory correctly abstracts the phenomenon, no empirical
experiment should falsify the predictions of that theory.

In \autocite{Herley2016unfalsifiability}, Cormac Herley explores what he calls
``an asymmetry in computer security'', which he defines as follows: ``Things
can be declared insecure by observation, but not the reverse. There is no
observation that allows us to declare an arbitrary system or technique
secure''. With security, Herley (as in the reminder of this paper) only 
focuses on cybersecurity and his intuition is that there is no scientific theory that can predict
the security of a system, nor a theory that can predict all possible
insecurities of a system (which, by negation may be used as a theory of
security).  Herley then uses this argument to show that ``claims that any
measure is necessary for security are empirically unfalsifiable''. An example
is whether or not password masking makes a system secure. Password masking
prevents a complete leakage of the password and a system in which password
masking is implemented is believed to be secure because an attacker won't be
seeing the password by ``shoulder surging''.  However, this reasoning is
circular and it is always true that a system in which password masking is
implemented is a system in which a password is masked. No experiment can ever
falsify this statement.  Given that any theory that is not falsifiable by an
empirical experiment is well known\footnote{``A theory which is not refutable
by any conceivable event is nonscientific. Irrefutability is not a virtue of a
theory (as people often think) but a vice.'' -- Karl Popper, Conjectures and
Refutations~\autocite{popper1962conjectures}.} to be nonscientific (i.e.
unfalsifiability is a fallacy of a theory), Herley concludes that there is no
scientific theory of cybersecurity; which means that cybersecurity lays in the
realm of pseudo-sciences~\autocite{Herley2016usenixvideo}. Herley, e.g.  in
\autocite{Herley2017justifying}, discusses the implications of a nonscientific
approach to cybersecurity, and highlights the tremendous impact on all the
scientific research and engineering of systems. While the criticism is
investigated in \autocite{Herley2016unfalsifiability}, no solution is provided
nor envisioned.

@@

\section{Literature Review}\label{sec:literature}
\begin{figure*}[t]
	\centering
	\includegraphics[width=.8\textwidth]{safety-security_3.pdf}
	\caption{Overview of keywords related to security and safety}
	\label{fig:safety-security}
\end{figure*}

In most of the natural languages the concepts of safety
and security are not syntactically differentiated and both terms (safety and
security) are expressed by the same word, e.g. sicurezza in Italian.  A
semantic distinction between safety and security is correlated to a
belief\footnote{A belief has to be intended as a proposition which is supposed
to be true by the majority of people in our society, without a scientific
underlying theory but based on partial empirical evidences or inductive reasoning based on partial empirical evidences.} that
safety deals with \emph{accidents} (i.e. an unfortunate incident) posed by the
natural environment (e.g. natural events such as wearing of hardware
components), whereas security deals with \emph{incidents} posed by mankind (e.g.
attackers and bugs). The fundamental difference between nature and mankind (and,
in turn, between safety and cybersecurity) is believed to be on the different
intents (incidents are intentional, whereas accidents are not) of the causes
that generates a threat. Namely, nature is believed not to have malicious
intents (but unfortunate causes-effects), whereas threats generated by mankind are
malicious\footnote{Of course, logical flaws or bugs may be
introduced by other means (e.g. ignorance) without explicit malicious intents,
but the exploitation of those flaws is considered (for now, and detailed
afterwards in the article) malicious, and thus we consider any vulnerability to
be malicious even if it is due to the lack of skills.}.
This conclusion seems to be against a general formulation of a cybersecurity theory;
how can we define a theory that predicts what a human being will do?
Cybersecurity attacks seem to be related to the creativity of the attacker
and thus unpredictable.

Currently, the most complete understanding of insecurity issues, is stored into
a network of databases of weaknesses (e.g.
CWE~\autocite{MITRE2020CWEresearch}), vulnerabilities (e.g. CVE~\autocite{CVE},
NVD~\autocite{NIST2020NVD}), and attacks (e.g. CAPEC~\autocite{MITRE2020CAPEC},
ATT\&CK~\autocite{MITRE2020ATTACK}).  Those insecurity issues can be related to
the violation of one or more requirements (explicit or implicit) in the
specification, design or implementation of a system. The correlation 
between insecurity flaws and security requirements has been used to define
standards such as the IEC 62443-1-3 (the Industrial communication networks -
Network and system security -- Part 3-3: System security requirements and
security levels) defines requirements as ``confidentiality of information
in transit/at-rest''. More generally, the idea of defining security requirements
as properties of a system was initially defined in 1970s with the 
CIA-triad (Confidentiality, Integrity, Availability) and refined 
over the decades introducing related concepts such as authenticity or
non-repudiation, or introducing new ones such as ``responsibility'' in the
RITE approach (see~\autocite{Samonas2014cia} for an overview of the evolution of the CIA-triad).

Systems are
nowadays tested against known or unknown attacks,
or verified against security requirements such as confidentiality.
However, no scientific falsifiable theory correlate
security requirements to necessary and sufficient conditions (e.g. mitigations) 
to declare a system secure.
